



Now add an analysis function for the Recycling Study Questions




# Recycling Study Questions RS01-RS06

* **Description:** These are questions included from a different study
    * **Coding Scheme:** 5-point agreement scale
   *  **Specific Items:**
       *  RS01: I generally don’t pay attention to how much energy I use.
       *  RS02: I would say I am very pro-environmental.
       *  RS03: I think saving energy is largely a waste of time.
        *  RS04: I am generally conservative on the political spectrum with regard to social issues.
        * RS05: I am generally conservative on the political spectrum with regard to economic issues.
        *  RS06: I consider myself knowledgeable about how much energy utilities use

head(rs,n=10)
# A tibble: 10 × 7
      id RS01                           RS02                           RS03                           RS04                           RS05                     RS06   
   <int> <dbl+lbl>                      <dbl+lbl>                      <dbl+lbl>                      <dbl+lbl>                      <dbl+lbl>                <dbl+l>
 1     1 2 [Somewhat Agree]             2 [Somewhat Agree]             5 [Disagree]                   4 [Somewhat Disagree]          4 [Somewhat Disagree]    4 [Som…
 2     2 5 [Disagree]                   3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 5 [Disagree]                   5 [Disagree]             3 [Nei…
 3     3 3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 3 [Neither agree nor di… 3 [Nei…
 4     4 4 [Somewhat Disagree]          3 [Neither agree nor disagree] 5 [Disagree]                   5 [Disagree]                   5 [Disagree]             3 [Nei…
 5     5 3 [Neither agree nor disagree] 2 [Somewhat Agree]             4 [Somewhat Disagree]          2 [Somewhat Agree]             3 [Neither agree nor di… 2 [Som…
 6     6 5 [Disagree]                   1 [Agree]                      5 [Disagree]                   5 [Disagree]                   5 [Disagree]             2 [Som…
 7     7 5 [Disagree]                   2 [Somewhat Agree]             5 [Disagree]                   5 [Disagree]                   2 [Somewhat Agree]       4 [Som…
 8     8 2 [Somewhat Agree]             3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 2 [Somewhat Agree]             1 [Agree]                3 [Nei…
 9     9 4 [Somewhat Disagree]          2 [Somewhat Agree]             5 [Disagree]                   5 [Disagree]                   5 [Disagree]             4 [Som…
10    10 5 [Disagree]                   1 [Agree]                      5 [Disagree]                   1 [Agree]                      1 [Agree]                1 [Agr…









```{r}
analyze_recycling_survey <- function(rs_data) {
  
  # 1) Coerce columns to numeric
  rs_numeric <- rs_data %>%
    mutate(
      RS01_num = as.numeric(as.character(RS01)),
      RS02_num = as.numeric(as.character(RS02)),
      RS03_num = as.numeric(as.character(RS03)),
      RS04_num = as.numeric(as.character(RS04)),
      RS05_num = as.numeric(as.character(RS05)),
      RS06_num = as.numeric(as.character(RS06))
    )
  
  # 2) Recode items so that higher numbers consistently reflect
  #    "more" of the targeted construct:
  #
  # Environmental Attitude (RS01 & RS03 are negative, RS02 & RS06 are positive).
  # Original scale is 1=Agree ... 5=Disagree
  # For a 'positive' pro-environment item, do 6 - x => so 1 => 5 (strong agreement => higher = pro-env).
  # For a 'negative' pro-environment item, keep x => so 1 => 1 (strong agreement => lower pro-env).
  
  rs_recode <- rs_numeric %>%
    mutate(
      # Positive items
      RS02_env = 6 - RS02_num,  # now 5 = strongly pro-env
      RS06_env = 6 - RS06_num,  # now 5 = strongly pro-env
      
      # Negative items (keep the original so that 1 => 1 = strongly anti-env, 5 => 5 = strongly pro-env)
      RS01_env = RS01_num,
      RS03_env = RS03_num,
      
      # Political items: If we want higher = more conservative,
      # we can do 6 - x so that 1 => 5 (strongly conservative).
      # If you'd prefer the raw code to remain 1=Agree => "lowest" numeric,
      # skip the transformation. Below we invert it:
      RS04_cons = 6 - RS04_num, 
      RS05_cons = 6 - RS05_num
    )
  
  # 3) Compute subscales
  #    - "env_attitude": average of RS01_env, RS02_env, RS03_env, RS06_env
  #      such that 5 = most pro-environment, 1 = least pro-environment
  #    - "pol_conservatism": average of RS04_cons, RS05_cons
  #      such that 5 = strongly conservative, 1 = strongly liberal
  
  rs_subscales <- rs_recode %>%
    rowwise() %>%
    mutate(
      env_attitude = mean(c(RS01_env, RS02_env, RS03_env, RS06_env), na.rm = TRUE),
      pol_conservatism = mean(c(RS04_cons, RS05_cons), na.rm = TRUE)
    ) %>%
    ungroup()
  
  # 4) Standardize subscales if desired
  rs_final <- rs_subscales %>%
    mutate(
      env_attitude_z = as.numeric(scale(env_attitude)),
      pol_conservatism_z = as.numeric(scale(pol_conservatism))
    ) %>%
    # 5) Return the columns of interest
    select(id, 
           env_attitude, env_attitude_z,
           pol_conservatism, pol_conservatism_z)
  
  return(rs_final)
}

rs_scores <- analyze_recycling_survey(rs)
head(rs_scores) |> pander::pandoc.table(caption = "Recycling Study Scores",split.table=Inf,style='rmarkdown')

```
| id | env_attitude | env_attitude_z | pol_conservatism | pol_conservatism_z |
|:--:|:------------:|:--------------:|:----------------:|:------------------:|
| 1  |     3.25     |    -0.4323     |        2         |       -0.449       |
| 2  |     3.5      |    -0.1079     |        1         |       -1.154       |
| 3  |      3       |    -0.7567     |        3         |       0.2555       |
| 4  |     3.75     |     0.2164     |        1         |       -1.154       |
| 5  |     3.75     |     0.2164     |       3.5        |       0.6077       |
| 6  |     4.75     |     1.514      |        1         |       -1.154       |

Table: Recycling Study Scores


Rationale for the Example Code
	•	RS02 (“I would say I am very pro-environmental”) and RS06 (“I consider myself knowledgeable …”) are phrased positively for environmental attitudes.
	•	If a person selects 1 [Agree], that is the strongest pro-environment response.
	•	Often, we prefer higher numbers to mean stronger agreement, so we do (6 - x) to flip the original code (1→5, 5→1).
	•	RS01 (“I generally don’t pay attention to how much energy I use”) and RS03 (“… saving energy is a waste of time”) are negative statements about the environment.
	•	If a person selects 1 [Agree] here, that is more anti-environment.
	•	To keep “5 = most pro-environment,” we just keep these as x in the final scale. (Hence, 1=least pro-environment, 5=most pro-environment for these reversed items.)
	•	Alternatively, one could unify them by also applying 6 - x but then interpret them as “bigger number = more negative.” The key is consistency.
	•	RS04 and RS05 ask about conservatism. One might keep these separate as a “political orientation” scale (higher = more conservative). For instance, if someone selects 1 [Agree], that means strongly conservative, so we do pol = 6 - x so that 5 indicates strong conservative and 1 indicates strong liberal.










```{r}

analyze_recycling_study <- function(rs) {
  # Reverse code items where higher values indicate disagreement with pro-environmental attitudes
  rs_recoded <- rs %>%
    mutate(
      RS01_rev = 6 - RS01,
      RS03_rev = 6 - RS03
    )
  
  # Calculate a composite score for pro-environmental attitudes (average of RS02, RS01_rev, RS03_rev)
  # and for political conservatism (average of RS04 and RS05)
  rs_scores <- rs_recoded %>%
    mutate(
      pro_env_attitude = rowMeans(select(., RS02, RS01_rev, RS03_rev), na.rm = TRUE),
      political_conservatism = rowMeans(select(., RS04, RS05), na.rm = TRUE),
      rs06_knowledge = RS06 # Rename RS06 to a more descriptive name
    )
  
  # Standardize the scores
  rs_standardized <- rs_scores %>%
    mutate(
      pro_env_attitude_std = scale(pro_env_attitude),
      political_conservatism_std = scale(political_conservatism),
      rs06_knowledge_std = scale(rs06_knowledge)
    )
  
  # Select relevant columns and return
  rs_final <- rs_standardized %>%
    select(id, pro_env_attitude, political_conservatism, rs06_knowledge,
           pro_env_attitude_std, political_conservatism_std, rs06_knowledge_std)
  
  return(rs_final)
}

# Example usage:
rs_scores <- analyze_recycling_study(rs)
head(rs_scores) |> pander::pandoc.table(caption = "Recycling Study Scores",split.table=Inf,style='rmarkdown')
```


| id | pro_env_attitude | political_conservatism | rs06_knowledge | pro_env_attitude_std | political_conservatism_std | rs06_knowledge_std |
|:--:|:----------------:|:----------------------:|:--------------:|:--------------------:|:--------------------------:|:------------------:|
| 1  |      2.333       |           4            |       4        |        0.1075        |           0.449            |       0.8629       |
| 2  |      2.333       |           5            |       3        |        0.1075        |           1.154            |      0.03801       |
| 3  |        3         |           3            |       3        |        0.8573        |          -0.2555           |      0.03801       |
| 4  |        2         |           5            |       3        |       -0.2674        |           1.154            |      0.03801       |
| 5  |      2.333       |          2.5           |       2        |        0.1075        |          -0.6077           |      -0.7869       |
| 6  |        1         |           5            |       2        |        -1.392        |           1.154            |      -0.7869       |

Table: Recycling Study Scores












```{r}


analyze_rs_survey <- function(rs) {
  # 1. Recode variables so higher scores = more pro-environmental
  rs_recoded <- rs %>%
    mutate(
      # Reverse code RS01 and RS03 (negative statements)
      RS01_rev = 6 - as.numeric(RS01),
      RS02_std = as.numeric(RS02),
      RS03_rev = 6 - as.numeric(RS03),
      # Political conservatism reverse coded as it typically correlates negatively with environmental concerns
      RS04_rev = 6 - as.numeric(RS04),
      RS05_rev = 6 - as.numeric(RS05),
      RS06_std = as.numeric(RS06)
    )
  
  # 2. Calculate subscores
  rs_scores <- rs_recoded %>%
    mutate(
      # Environmental attitude subscore (RS01, RS02, RS03)
      env_attitude = (RS01_rev + RS02_std + RS03_rev)/3,
      
      # Political orientation subscore (RS04, RS05)
      political_orientation = (RS04_rev + RS05_rev)/2,
      
      # Knowledge score (RS06)
      env_knowledge = RS06_std
    )
  
  # 3. Calculate composite score with weights
  final_scores <- rs_scores %>%
    mutate(
      # Weight environmental attitude more heavily
      rs_composite = (0.5 * env_attitude + 
                     0.3 * political_orientation +
                     0.2 * env_knowledge),
      
      # Standardized versions
      env_attitude_z = scale(env_attitude)[,1],
      political_orientation_z = scale(political_orientation)[,1],
      env_knowledge_z = scale(env_knowledge)[,1],
      rs_composite_z = scale(rs_composite)[,1]
    ) %>%
    select(id, 
           env_attitude, env_attitude_z,
           political_orientation, political_orientation_z,
           env_knowledge, env_knowledge_z,
           rs_composite, rs_composite_z)
  
  return(final_scores)
}

rs_scores <- analyze_rs_survey(rs) |> head(5) |> pander::pandoc.table(caption = "Recycling Study Scores",split.table=Inf,style='rmarkdown')

```

| id | env_attitude | env_attitude_z | political_orientation | political_orientation_z | env_knowledge | env_knowledge_z | rs_composite | rs_composite_z |
|:--:|:------------:|:--------------:|:---------------------:|:-----------------------:|:-------------:|:---------------:|:------------:|:--------------:|
| 1  |    2.333     |     0.1075     |           2           |         -0.449          |       4       |     0.8629      |    2.567     |    0.08558     |
| 2  |    2.333     |     0.1075     |           1           |         -1.154          |       3       |     0.03801     |    2.067     |    -0.5649     |
| 3  |      3       |     0.8573     |           3           |         0.2555          |       3       |     0.03801     |      3       |     0.6493     |
| 4  |      2       |    -0.2674     |           1           |         -1.154          |       3       |     0.03801     |     1.9      |    -0.7817     |
| 5  |    2.333     |     0.1075     |          3.5          |         0.6077          |       2       |     -0.7869     |    2.617     |     0.1506     |

Table: Recycling Study Scores


his analysis approach:

Handles reverse coding for negatively worded items (RS01, RS03)
Creates meaningful subscores:
    Environmental attitude (RS01-RS03)
    Political orientation (RS04-RS05)
    Environmental knowledge (RS06)
Weights the components based on their relevance to environmental behavior:
    50% environmental attitude
    30% political orientation
    20% environmental knowledge
Provides both raw and standardized scores
Includes a combination function to merge with RECS scores for a comprehensive assessment
The combination function allows us to:

Compare behavioral measures (RECS) with attitudinal measures (RS)
Create an overall environmental score
Identify potential gaps between attitudes and behaviors