



Now add an analysis function for the Recycling Study Questions




# Recycling Study Questions RS01-RS06

* **Description:** These are questions included from a different study
    * **Coding Scheme:** 5-point agreement scale
   *  **Specific Items:**
       *  RS01: I generally don’t pay attention to how much energy I use.
       *  RS02: I would say I am very pro-environmental.
       *  RS03: I think saving energy is largely a waste of time.
        *  RS04: I am generally conservative on the political spectrum with regard to social issues.
        * RS05: I am generally conservative on the political spectrum with regard to economic issues.
        *  RS06: I consider myself knowledgeable about how much energy utilities use

head(rs,n=10)
# A tibble: 10 × 7
      id RS01                           RS02                           RS03                           RS04                           RS05                     RS06   
   <int> <dbl+lbl>                      <dbl+lbl>                      <dbl+lbl>                      <dbl+lbl>                      <dbl+lbl>                <dbl+l>
 1     1 2 [Somewhat Agree]             2 [Somewhat Agree]             5 [Disagree]                   4 [Somewhat Disagree]          4 [Somewhat Disagree]    4 [Som…
 2     2 5 [Disagree]                   3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 5 [Disagree]                   5 [Disagree]             3 [Nei…
 3     3 3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 3 [Neither agree nor di… 3 [Nei…
 4     4 4 [Somewhat Disagree]          3 [Neither agree nor disagree] 5 [Disagree]                   5 [Disagree]                   5 [Disagree]             3 [Nei…
 5     5 3 [Neither agree nor disagree] 2 [Somewhat Agree]             4 [Somewhat Disagree]          2 [Somewhat Agree]             3 [Neither agree nor di… 2 [Som…
 6     6 5 [Disagree]                   1 [Agree]                      5 [Disagree]                   5 [Disagree]                   5 [Disagree]             2 [Som…
 7     7 5 [Disagree]                   2 [Somewhat Agree]             5 [Disagree]                   5 [Disagree]                   2 [Somewhat Agree]       4 [Som…
 8     8 2 [Somewhat Agree]             3 [Neither agree nor disagree] 3 [Neither agree nor disagree] 2 [Somewhat Agree]             1 [Agree]                3 [Nei…
 9     9 4 [Somewhat Disagree]          2 [Somewhat Agree]             5 [Disagree]                   5 [Disagree]                   5 [Disagree]             4 [Som…
10    10 5 [Disagree]                   1 [Agree]                      5 [Disagree]                   1 [Agree]                      1 [Agree]                1 [Agr…









```{r}
analyze_recycling_survey <- function(rs_data) {
  
  # 1) Coerce columns to numeric
  rs_numeric <- rs_data %>%
    mutate(
      RS01_num = as.numeric(as.character(RS01)),
      RS02_num = as.numeric(as.character(RS02)),
      RS03_num = as.numeric(as.character(RS03)),
      RS04_num = as.numeric(as.character(RS04)),
      RS05_num = as.numeric(as.character(RS05)),
      RS06_num = as.numeric(as.character(RS06))
    )
  
  # 2) Recode items so that higher numbers consistently reflect
  #    "more" of the targeted construct:
  #
  # Environmental Attitude (RS01 & RS03 are negative, RS02 & RS06 are positive).
  # Original scale is 1=Agree ... 5=Disagree
  # For a 'positive' pro-environment item, do 6 - x => so 1 => 5 (strong agreement => higher = pro-env).
  # For a 'negative' pro-environment item, keep x => so 1 => 1 (strong agreement => lower pro-env).
  
  rs_recode <- rs_numeric %>%
    mutate(
      # Positive items
      RS02_env = 6 - RS02_num,  # now 5 = strongly pro-env
      RS06_env = 6 - RS06_num,  # now 5 = strongly pro-env
      
      # Negative items (keep the original so that 1 => 1 = strongly anti-env, 5 => 5 = strongly pro-env)
      RS01_env = RS01_num,
      RS03_env = RS03_num,
      
      # Political items: If we want higher = more conservative,
      # we can do 6 - x so that 1 => 5 (strongly conservative).
      # If you'd prefer the raw code to remain 1=Agree => "lowest" numeric,
      # skip the transformation. Below we invert it:
      RS04_cons = 6 - RS04_num, 
      RS05_cons = 6 - RS05_num
    )
  
  # 3) Compute subscales
  #    - "env_attitude": average of RS01_env, RS02_env, RS03_env, RS06_env
  #      such that 5 = most pro-environment, 1 = least pro-environment
  #    - "pol_conservatism": average of RS04_cons, RS05_cons
  #      such that 5 = strongly conservative, 1 = strongly liberal
  
  rs_subscales <- rs_recode %>%
    rowwise() %>%
    mutate(
      env_attitude = mean(c(RS01_env, RS02_env, RS03_env, RS06_env), na.rm = TRUE),
      pol_conservatism = mean(c(RS04_cons, RS05_cons), na.rm = TRUE)
    ) %>%
    ungroup()
  
  # 4) Standardize subscales if desired
  rs_final <- rs_subscales %>%
    mutate(
      env_attitude_z = as.numeric(scale(env_attitude)),
      pol_conservatism_z = as.numeric(scale(pol_conservatism))
    ) %>%
    # 5) Return the columns of interest
    select(id, 
           env_attitude, env_attitude_z,
           pol_conservatism, pol_conservatism_z)
  
  return(rs_final)
}

rs_scores <- analyze_recycling_survey(rs)
head(rs_scores) |> pander::pandoc.table(caption = "Recycling Study Scores",split.table=Inf,style='rmarkdown')

```
| id | env_attitude | env_attitude_z | pol_conservatism | pol_conservatism_z |
|:--:|:------------:|:--------------:|:----------------:|:------------------:|
| 1  |     3.25     |    -0.4323     |        2         |       -0.449       |
| 2  |     3.5      |    -0.1079     |        1         |       -1.154       |
| 3  |      3       |    -0.7567     |        3         |       0.2555       |
| 4  |     3.75     |     0.2164     |        1         |       -1.154       |
| 5  |     3.75     |     0.2164     |       3.5        |       0.6077       |
| 6  |     4.75     |     1.514      |        1         |       -1.154       |

Table: Recycling Study Scores


Rationale for the Example Code
	•	RS02 (“I would say I am very pro-environmental”) and RS06 (“I consider myself knowledgeable …”) are phrased positively for environmental attitudes.
	•	If a person selects 1 [Agree], that is the strongest pro-environment response.
	•	Often, we prefer higher numbers to mean stronger agreement, so we do (6 - x) to flip the original code (1→5, 5→1).
	•	RS01 (“I generally don’t pay attention to how much energy I use”) and RS03 (“… saving energy is a waste of time”) are negative statements about the environment.
	•	If a person selects 1 [Agree] here, that is more anti-environment.
	•	To keep “5 = most pro-environment,” we just keep these as x in the final scale. (Hence, 1=least pro-environment, 5=most pro-environment for these reversed items.)
	•	Alternatively, one could unify them by also applying 6 - x but then interpret them as “bigger number = more negative.” The key is consistency.
	•	RS04 and RS05 ask about conservatism. One might keep these separate as a “political orientation” scale (higher = more conservative). For instance, if someone selects 1 [Agree], that means strongly conservative, so we do pol = 6 - x so that 5 indicates strong conservative and 1 indicates strong liberal.




ea_items_rev <- rs %>% select(RS01:RS06) #re-select items
alpha_ea_rev <- psych::alpha(ea_items_rev, check.keys=TRUE)
print("Cronbach's Alpha for Environmental Attitude Scale (check.keys=TRUE):")
print(alpha_ea_rev)
[1] "Cronbach's Alpha for Environmental Attitude Scale (check.keys=TRUE):"

Reliability analysis   
Call: psych::alpha(x = ea_items_rev, check.keys = TRUE)

  raw_alpha std.alpha G6(smc) average_r S/N   ase mean   sd median_r
      0.65      0.64    0.71      0.23 1.8 0.022  3.5 0.79     0.21

    95% confidence boundaries 
         lower alpha upper
Feldt     0.61  0.65  0.69
Duhachek  0.61  0.65  0.69

 Reliability if an item is dropped:
      raw_alpha std.alpha G6(smc) average_r S/N alpha se var.r med.r
RS01       0.59      0.57    0.64      0.21 1.3    0.025 0.068  0.24
RS02-      0.61      0.59    0.68      0.22 1.4    0.025 0.079  0.21
RS03       0.60      0.59    0.64      0.22 1.4    0.026 0.052  0.21
RS04       0.56      0.56    0.59      0.21 1.3    0.029 0.036  0.24
RS05       0.54      0.55    0.58      0.20 1.2    0.030 0.039  0.21
RS06-      0.72      0.71    0.74      0.33 2.5    0.018 0.032  0.27

 Item statistics 
        n raw.r std.r r.cor r.drop mean  sd
RS01  586  0.64  0.66  0.57  0.423  3.4 1.3
RS02- 586  0.59  0.62  0.47  0.396  3.6 1.1
RS03  586  0.60  0.63  0.54  0.424  4.3 1.1
RS04  586  0.72  0.67  0.66  0.509  3.5 1.5
RS05  586  0.74  0.69  0.68  0.531  3.2 1.5
RS06- 586  0.29  0.32  0.11  0.038  3.0 1.2

Non missing response frequency for each item
        1    2    3    4    5 miss
RS01 0.09 0.24 0.10 0.31 0.26    0
RS02 0.25 0.34 0.24 0.12 0.05    0
RS03 0.03 0.08 0.09 0.23 0.58    0
RS04 0.14 0.17 0.17 0.13 0.40    0
RS05 0.18 0.21 0.14 0.12 0.35    0
RS06 0.10 0.32 0.23 0.22 0.13    0




```{r}

analyze_recycling_study <- function(rs) {
  # Reverse code items where higher values indicate disagreement with pro-environmental attitudes
  rs_recoded <- rs %>%
    mutate(
      RS01_rev = 6 - RS01,
      RS03_rev = 6 - RS03
    )
  
  # Calculate a composite score for pro-environmental attitudes (average of RS02, RS01_rev, RS03_rev)
  # and for political conservatism (average of RS04 and RS05)
  rs_scores <- rs_recoded %>%
    mutate(
      pro_env_attitude = rowMeans(select(., RS02, RS01_rev, RS03_rev), na.rm = TRUE),
      political_conservatism = rowMeans(select(., RS04, RS05), na.rm = TRUE),
      rs06_knowledge = RS06 # Rename RS06 to a more descriptive name
    )
  
  # Standardize the scores
  rs_standardized <- rs_scores %>%
    mutate(
      pro_env_attitude_std = scale(pro_env_attitude),
      political_conservatism_std = scale(political_conservatism),
      rs06_knowledge_std = scale(rs06_knowledge)
    )
  
  # Select relevant columns and return
  rs_final <- rs_standardized %>%
    select(id, pro_env_attitude, political_conservatism, rs06_knowledge,
           pro_env_attitude_std, political_conservatism_std, rs06_knowledge_std)
  
  return(rs_final)
}

# Example usage:
rs_scores <- analyze_recycling_study(rs)
head(rs_scores) |> pander::pandoc.table(caption = "Recycling Study Scores",split.table=Inf,style='rmarkdown')
```


| id | pro_env_attitude | political_conservatism | rs06_knowledge | pro_env_attitude_std | political_conservatism_std | rs06_knowledge_std |
|:--:|:----------------:|:----------------------:|:--------------:|:--------------------:|:--------------------------:|:------------------:|
| 1  |      2.333       |           4            |       4        |        0.1075        |           0.449            |       0.8629       |
| 2  |      2.333       |           5            |       3        |        0.1075        |           1.154            |      0.03801       |
| 3  |        3         |           3            |       3        |        0.8573        |          -0.2555           |      0.03801       |
| 4  |        2         |           5            |       3        |       -0.2674        |           1.154            |      0.03801       |
| 5  |      2.333       |          2.5           |       2        |        0.1075        |          -0.6077           |      -0.7869       |
| 6  |        1         |           5            |       2        |        -1.392        |           1.154            |      -0.7869       |

Table: Recycling Study Scores
















```{r}


analyze_rs_survey <- function(rs) {
  # 1. Recode variables so higher scores = more pro-environmental
  rs_recoded <- rs %>%
    mutate(
      # Reverse code RS01 and RS03 (negative statements)
      RS01_rev = 6 - as.numeric(RS01),
      RS02_std = as.numeric(RS02),
      RS03_rev = 6 - as.numeric(RS03),
      # Political conservatism reverse coded as it typically correlates negatively with environmental concerns
      RS04_rev = 6 - as.numeric(RS04),
      RS05_rev = 6 - as.numeric(RS05),
      RS06_std = as.numeric(RS06)
    )
  
  # 2. Calculate subscores
  rs_scores <- rs_recoded %>%
    mutate(
      # Environmental attitude subscore (RS01, RS02, RS03)
      env_attitude = (RS01_rev + RS02_std + RS03_rev)/3,
      
      # Political orientation subscore (RS04, RS05)
      political_orientation = (RS04_rev + RS05_rev)/2,
      
      # Knowledge score (RS06)
      env_knowledge = RS06_std
    )
  
  # 3. Calculate composite score with weights
  final_scores <- rs_scores %>%
    mutate(
      # Weight environmental attitude more heavily
      rs_composite = (0.5 * env_attitude + 
                     0.3 * political_orientation +
                     0.2 * env_knowledge),
      
      # Standardized versions
      env_attitude_z = scale(env_attitude)[,1],
      political_orientation_z = scale(political_orientation)[,1],
      env_knowledge_z = scale(env_knowledge)[,1],
      rs_composite_z = scale(rs_composite)[,1]
    ) %>%
    select(id, 
           env_attitude, env_attitude_z,
           political_orientation, political_orientation_z,
           env_knowledge, env_knowledge_z,
           rs_composite, rs_composite_z)
  
  return(final_scores)
}

rs_scores <- analyze_rs_survey(rs) |> head(5) |> pander::pandoc.table(caption = "Recycling Study Scores",split.table=Inf,style='rmarkdown')

```

| id | env_attitude | env_attitude_z | political_orientation | political_orientation_z | env_knowledge | env_knowledge_z | rs_composite | rs_composite_z |
|:--:|:------------:|:--------------:|:---------------------:|:-----------------------:|:-------------:|:---------------:|:------------:|:--------------:|
| 1  |    2.333     |     0.1075     |           2           |         -0.449          |       4       |     0.8629      |    2.567     |    0.08558     |
| 2  |    2.333     |     0.1075     |           1           |         -1.154          |       3       |     0.03801     |    2.067     |    -0.5649     |
| 3  |      3       |     0.8573     |           3           |         0.2555          |       3       |     0.03801     |      3       |     0.6493     |
| 4  |      2       |    -0.2674     |           1           |         -1.154          |       3       |     0.03801     |     1.9      |    -0.7817     |
| 5  |    2.333     |     0.1075     |          3.5          |         0.6077          |       2       |     -0.7869     |    2.617     |     0.1506     |

Table: Recycling Study Scores


his analysis approach:

Handles reverse coding for negatively worded items (RS01, RS03)
Creates meaningful subscores:
    Environmental attitude (RS01-RS03)
    Political orientation (RS04-RS05)
    Environmental knowledge (RS06)
Weights the components based on their relevance to environmental behavior:
    50% environmental attitude
    30% political orientation
    20% environmental knowledge
Provides both raw and standardized scores
Includes a combination function to merge with RECS scores for a comprehensive assessment
The combination function allows us to:

Compare behavioral measures (RECS) with attitudinal measures (RS)
Create an overall environmental score
Identify potential gaps between attitudes and behaviors